---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  dev = "svglite"
)
```

# loco

<!-- badges: start -->
<!-- badges: end -->

This package implements the calculation of local correlation (LoCo) similarity scores for a pair of
time series, proposed by Papadimitriou, S., Sun, J., and Yu, P.S., in the paper
["Local Correlation Tracking in Time Series"](https://doi.org/10.1109/ICDM.2006.99).

## Installation

You can install this package using:

``` r
# install.packages("remotes")
remotes::install_github("adamoshen/loco")
```

## Basic usage

```{r}
library(tidyverse)
library(patchwork)
library(loco)
```

Suppose we are interested in the local correlations of average daily temperature readings at DTW
and LGA airports. A preview of the data:

```{r}
daily_temperature %>%
  group_by(origin) %>%
  slice_head(n = 6) %>%
  group_split()
```

First, we require a data frame containing the two time series' data as columns.

```{r}
daily_temperature_wide <- daily_temperature %>%
  pivot_wider(id_cols=date, names_from=origin, values_from=temp)

daily_temperature_wide
```

We should also double-check that our two time series are aligned, i.e. that there are no missing
values in either series. Treatment of missing values is left to the user.

```{r}
daily_temperature_wide %>%
  summarise(
    missing_dtw = any(is.na(DTW)),
    missing_lga = any(is.na(LGA))
  )
```

The LoCo scores can be obtained by passing `daily_temperature_wide` to `loco::loco()`. We set the
window size to 5 and the number of principal eigenvectors used to 3. We can optionally supply the
`date` column to the `timestamps` argument for ease of subsequent visualisation.

```{r}
loco_scores <- daily_temperature_wide %>%
  loco(DTW, LGA, timestamps=date, window_size=5, k=3)

loco_scores
```

We can now plot the LoCo scores and identify time points of particular interest.

```{r}
p1 <- ggplot(daily_temperature_wide) +
  geom_line(aes(x=date, y=DTW), colour="#009E73", alpha=0.6) +
  geom_line(aes(x=date, y=LGA), colour="#D55E00", alpha=0.6) +
  scale_x_date(date_breaks="2 months", date_labels="%b") +
  labs(x="", y="Average daily temperature (fahrenheit)") +
  theme_bw()

p2 <- ggplot(loco_scores) +
  geom_line(aes(x=timestamps, y=scores), colour="black") +
  scale_x_date(date_breaks="2 months", date_labels="%b") +
  labs(x="", y="LoCo score") +
  theme_bw()

p1 / p2
```
